

%%% Choose a language %%%

\newif\ifEN
\ENtrue   % uncomment this for english
%\ENfalse   % uncomment this for czech

%%% Configuration of the title page %%%

\newif\ifMFF
\MFFtrue % comment this out for the version with a big UK university logo
\def\UKName{Charles University in Prague} %this is only used in UK-logo-version
\def\UKFaculty{Faculty of Mathematics and Physics}

% Thesis type names, as used in several places in the title
%\def\ThesisTypeTitle{\ifEN BACHELOR THESIS \else BAKALÁŘSKÁ PRÁCE \fi}
\def\ThesisTypeTitle{\ifEN MASTER THESIS \else DIPLOMOVÁ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN RIGOROUS THESIS \else RIGORÓZNÍ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN DOCTORAL THESIS \else DISERTAČNÍ PRÁCE \fi}
%\%def\ThesisGenitive{\ifEN bachelor \else bakalářské \fi}
\def\ThesisGenitive{\ifEN master \else diplomové \fi}
%\def\ThesisGenitive{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisGenitive{\ifEN doctoral \else disertační \fi}
%\def\ThesisAccusative{\ifEN bachelor \else bakalářskou \fi}
\def\ThesisAccusative{\ifEN master \else diplomovou \fi}
%\def\ThesisAccusative{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisAccusative{\ifEN doctoral \else disertační \fi}



%%% Fill in your details %%%

% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{Prediction of velocity and speed of movement from human intracranial EEG recordings using deep neural networks}
\def\ThesisAuthor{Michaela Vystrčilová}
\def\YearSubmitted{2021}

% department assigned to the thesis
\def\Department{Department of Software and Computer Science Education}
% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

\def\Supervisor{Mgr. Ján Antolík, PhD}
\def\SupervisorsDepartment{Department of Software and Computer Science Education}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{IUI}

\def\Dedication{First, I would like to thank my supervisor Mgr. Ján Antolík, Ph.D. for introducing me to the intriguing field of neuroscience. His supportive attitude and advice helped me way beyond this thesis. In a similar manner I would like to thank Mgr. Jiří Hammer, Ph.D. who offered me not only this project but importantly also his time and knowledge. 

I also want to express my gratitude to Mr. Robin Tibor Schirrmeister, soon to be Ph.D. and Prof. Tonio Ball for their invaluable and welcoming consultations regarding the Deep4Net and brain signal processing in general.

Further I would like to manifest my appreciation for at least some of the people who, more or less directly, contributed to me starting and also finishing this thesis. For encouraging me not to content myself with something I would not enjoy doing, for standing me in times of quarantine and for all the emotional support and food, thank you, Vít Kabele, Jana, Jiří, Jana and Lenka Vystrčilovi, Anna Bláhová. 

Last, but not least, I would like to thank my cat for sitting on my lap long enough for me to be able to finish the thesis. 
}

\def\AbstractEN{%
Our brain controls the processes of the body including movement. In this thesis, we try to understand how the information about hand movement is encoded into the brain's electrical activity and how this activity can be used to predict the velocity and absolute velocity of hand movements. Using a well-established deep neural network architecture for EEG decoding - the Deep4Net - we predict hand movement velocity and absolute velocity from intracranial EEG signals. While reaching the expected performance level, we determine the influence of different frequency bands on the network's prediction. We find that modulations in the high-gamma frequency band are less informative than expected based on previous studies. We also identify two architectural modifications which lead to higher performances. 1. the removal of max-pooling layers in the architecture leads to significantly higher correlations. 2. the non-uniform receptive field of the network is a potential drawback making the network biased towards less relevant information. 
% ABSTRACT IS NOT A COPY OF YOUR THESIS ASSIGNMENT!
}

\def\AbstractCS{Lidský mozek řídí naše těla včetně pohybů rukou. V této diplomové práci jsme se snažili porozumět tomu, jak je informace o pohybu rukou zakódovaná do elektrické aktivity mozku a jak lze tuto aktivitu použít při predikování rychlosti a absolutní rychlosti pohybu rukou. Pomocí osvědčené hluboké neuronové sítě (Deep4Net), používané k dekódování signálů z EEG, jsme dekódovali rychlost a absolutní rychlost pohybů rukou z intrakraniálního EEG. Úspěšnějsme dosáhli očekávaných hodnot korelace mezi predikovanými a skutečnými proměnnými. Zároveň jsme stanovili vliv amplitud vlnových pásem v různých frekveních na predikce této sítě. Zjistili jsme, že vliv modulací ve vlnovém pásmu high-gamma je menší než se očekávalo na základě předchozích studií. Současně jsme identifikovali dvě možné úpravy architektury Deep4Net, které mohou vést k lepším predikčním schopnostem.  Zaprvé jsme odhalili, že odstraněním max-poolingových vrstev lze dosáhnout statisticky signifikantně lepších korelačních koeficientů. Zadruhé jsme objevili, že nerovnoměrné receptivní pole této sítě představuje nevýhodu a způsobuje že se vícesoustřeďují na méně relevantní informace.}

% 3 to 5 keywords (recommended), each enclosed in curly braces.
% Keywords are useful for indexing and searching for the theses by topic.
\def\Keywords{%
{{motor decoding}, {intracranial EEG (iEEG)}, {deep learning}, {convolutional neural networks}}
}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
% Alternatively, consider increasing the size of the page by uncommenting the
% geometry modification in thesis.tex.
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  %uncomment to decrease font size

\ifEN\relax\else
% If you are writing a czech thesis, you additionally need to fill in the
% english translation of the metadata here!
\def\ThesisTitleEN{\xxx{Thesis title in English}}
\def\DepartmentEN{\xxx{Name of the department in English}}
\def\DeptTypeEN{\xxx{Department}}
\def\SupervisorsDepartmentEN{\xxx{Superdepartment}}
\def\StudyProgrammeEN{\xxx{study programme}}
\def\StudyBranchEN{\xxx{study branch}}
\def\KeywordsEN{%
\xxx{{key} {words}}
}
\fi
